---
title: "Modelling"
author: "Mohsen Bahremani"
date: "11/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
library(dplyr)
library(ggplot2)
library(scales)
library(readr)
library(tidyverse)
library(modelr)
library(broom)
library(knitr)
```
```{r}
url='https://ed-public-download.app.cloud.gov/downloads/Most-Recent-Cohorts-All-Data-Elements.csv'
df <- read_csv(url)
```
There are 7593 colleges in this dataset with 1777 variables. For this analysis, we are only interested in several defining characteristics (e.g. institution name), categorical factors (e.g. operational status), and quantitative variables (e.g. admission rate).

The following code selects only the variables relevant to our analysis and only retains colleges that:

    are currently operating
    are in the 50 states (i.e. excluding territories)
    are not for-profit institutions
    predominantly award bachelor’s degrees

```{r}
college <- df %>%
  filter(CURROPER == 1,
         ST_FIPS <= 56,
         CONTROL != 3,
         PREDDEG == 3) %>%
  select(name = INSTNM, funding = CONTROL, admit = ADM_RATE, med_earnings = MD_EARN_WNE_P10,
         med_fam_inc = MD_FAMINC, NPT4_PUB, NPT4_PRIV)
```

In order to create visualizations of the data, we need to convert the variables admit, med_earnings, med_fam_inc, NPT4_PUB, and NPT4_PRIV from the character type to the double type. The following code also merges the NPT4_PUB column and NPT4_PRIV column into a single price column.

```{r}
college$admit <- as.double(college$admit)
college$med_earnings <- as.double(college$med_earnings)
college$med_fam_inc <- as.double(college$med_fam_inc)
college$NPT4_PUB <- as.double(college$NPT4_PUB)
college$NPT4_PRIV <- as.double(college$NPT4_PRIV)

college <- college %>%
  rowwise %>%
  mutate(price = sum(NPT4_PUB, NPT4_PRIV, na.rm = TRUE)) %>%
  select(-c(NPT4_PUB, NPT4_PRIV))

college["price"][college["price"] == 0] <- NA
```

```{r}
head(college)
```
```{r}
names(college)
```


1- funding refers to the source of funding for the institution, 1 coding for a public university and 2 coding for a private university
2-  admit is the admission rate of the institution on a scale of 0 to 1
3-  med_earnings represents the median earnings of the institution’s students who are employed 10 years after enrollment in 2015 USD
4-  med_fam_inc is the median family income of the institution’s current students in 2015 USD
5-  price indicates the average net price of attendance in USD accounting for the full costs of attendance and awarded financial aid

The following code stores certain purely cosmetic alterations of the visualizations as the variables xdollar, ydollar, and titling to allow for cleaner looking code.

```{r}
xdollar <- c(scale_x_continuous(labels = dollar,
                              breaks = seq(0, 130000, 25000),
                              limits = c(0, NA)))

ydollar <- c(scale_y_continuous(labels = dollar,
                                        breaks = seq(0, 130000, 25000),
                                        limits = c(0, NA)))

titling <- theme(plot.title = element_text(hjust = 0.5,
                                           face = "bold"),
                 axis.title.x = element_text(face = "bold"),
                 axis.title.y = element_text(face = "bold"))
```
##Distribution of Median Earnings
To analyze the relationship between median earnings and other factors, we would first like to get a preliminary understanding of the distribution of the median earnings of colleges’ graduates. We will create a boxplot of med_earnings below:
```{r}
ggplot(data = college) +
  geom_boxplot(mapping = aes(x = "", y = med_earnings)) +
  labs(title = "Median Earnings of \na College's Graduates",
       x = NULL,
       y = "Median Earnings in USD") +
  ydollar +
  titling
```
Due to the inclusion of outliers in our boxplot, we do not receive a good representation of the scale of the distribution of median earnings. Let’s take a look at a histogram instead:
```{r}
ggplot(data = college) +
  geom_histogram(mapping = aes(x = med_earnings)) +
  labs(title = "Median Earnings of a College's Graduates",
       x = "Median Earnings in USD",
       y = "Frequency of Colleges") +
  xdollar +
  titling
```

##Differences between the Median Earnings of Public Universities versus Private Universities
A histogram of med_earnings gives us a better visualization of the distribution of the median earnings of colleges’ graduates. However, to better understand the effect of different characteristics on economic outcomes, we would like to separate and compare the distributions of med_earnings between public and private institutions.

A violin plot should combine the compactness of a boxplot with the visualization of the distribution of a histogram. Additionally, I have overlaid a plot of the data points to further aid with visualizing the distribution. Here is a violin plot of med_earnings separated by the two values of funding:

```{r}
ggplot(data = college,
       mapping = aes(x = factor(funding),
                     y = med_earnings)) +
  geom_violin() +
  geom_jitter(alpha = 0.15) +
  scale_x_discrete(labels = c("Public", "Private")) +
  labs(title = "Median Earnings of a College's Graduates \nby Source of Funding",
       x = "Funding Source",
       y = "Median Earnings in USD") +
  ydollar +
  titling
```
Constructing a violin plot does well in illustrating the visible difference in the distributions of med_earnings between different kinds of universities. However, we could verify the difference in distribution computationally by a Kolmogorov-Smirnov test.

##Model
In order to further examine the effect various factors have on the economic outcome of a college’s graduates, it would serve us well to fit a linear model to the data. Constructing a least squares regression line using med_earnings and another variable will allow us to quantitatively observe the two variables’ relationship and determine the strength of the relationship, which might allow us to establish which variables are better predictors of good economic incomes than others.
```{r}
pairs(~med_earnings+price+log(admit)+log(med_fam_inc),
      data=college, lower.panel=panel.smooth)
```

```{r}
model.fit<- lm(med_earnings~price+admit+med_fam_inc, data=college)
summary(model.fit)
```
The Multiple R-squared value of 0.3181 means that 31.8% of the variation in median earnings can be explained by the linear relationship between median earnings and admission rate. This is a rather low value for R2. It is possible that a linear model is not the best fit for the data.  

```{r}
par(mfrow = c(1,2))
plot(model.fit,1)
plot(model.fit,2)

```
The model diagnostic plots above show that the model is passable. There is good scatter of the residuals around zero for the range of fitted values (the mean value of the residuals is, in fact, zero). The residuals Q-Q plot shows a pretty normal distribution. Overall, the evidence points toward the final model being valid.

```{r}

```

```{r}
library(car)
a= outlierTest(model.fit)
a
```


```{r}
college2 <- college[-c(779,613,878,615,1627,605,540,1289),]
model.fit2<- lm(med_earnings~price+admit+med_fam_inc, data=college2)
summary(model.fit2)
```
R_squared became 0.3514, and show us removing outliers help us to have better model. 

