---
title: "pipe review"
author: "Mohsen Bahremani"
date: "10/18/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(gapminder)

```

`dplyr` uses the forward pipe operator, usually called simply a **pipe**. We write pipes like `%>%` (`Ctrl+Shift+M`), you can read out loud: "and then..."

Pipes take the object on the *left* and apply the function on the *right*: `x %>% f(y) = f(x, y)`.

```{r}
gapminder %>% 
  filter(country == "Canada") %>% 
  head(2)
```

## data 
Use `.` in the data argument whenever you find yourself piping a function where data are not the first argument.

```{r}
gapminder %>%
  filter(country == "Canada") %>% 
  lm(pop ~ year, data = .)
```

We can use `%in%` like `==` but for matching *any element* in the vector on its right<sup>1</sup>. 

```{r}
former_yugoslavia <- c("Bosnia and Herzegovina", "Croatia", 
                       "Montenegro", "Serbia", "Slovenia")
yugoslavia <- gapminder %>% 
                  filter(country %in% former_yugoslavia)
yugoslavia
```

## `distinct()`

You can see all the *unique values* in your data for combinations of columns using `distinct()`:
```{r}
gapminder %>% 
    distinct(continent, year)

```
 if we need other features(variables) when we are using distinct:
```{r}
gapminder %>% distinct(continent, .keep_all=TRUE)
```
## Sorting: `arrange()`

Along with filtering the data to see certain rows, we might want to sort it:
The data are sorted by ascending `year` and descending `pop`.


```{r}
yugoslavia %>% 
  arrange(year, desc(pop))
```
## Keeping Columns: `select()`

Not only can we limit rows, but we can include specific columns (and put them in the order listed) using `select()`. 
```{r}
yugoslavia %>% 
  select(country, year, pop) %>% 
  head(4)

```
## Dropping Columns: `select()`


We can instead drop only specific columns with `select()` using `-` signs:
```{r}
yugoslavia %>% 
  select(-continent, -pop, -lifeExp) %>% 
  head(4)
```


## Renaming Columns with `select()`


We can rename columns using `select()`, but that drops everything that isn't mentioned:

```{r}
yugoslavia %>%
    select(Life_Expectancy = lifeExp) %>%
    head(4)
```


## Create New Columns: `mutate()`

In `dplyr`, you can add new columns to a data frame using `mutate()`.

```{r}
yugoslavia %>% filter(country == "Serbia") %>%
    select(year, pop, lifeExp) %>%
    mutate(pop_million = pop / 1000000,
           life_exp_past_40 = lifeExp - 40) %>%
    head(5)
```


## General Aggregation: `summarize()`

`summarize()` takes your column(s) of data and computes something using every row:

* Count how many rows there are
* Calculate the mean
* Compute the sum
* Obtain a minimum or maximum value
* ...

Useful functions
    Center: mean(), median()
    Spread: sd(), IQR(), mad()
    Range: min(), max(), quantile()
    Position: first(), last(), nth()
    Count: n(), n_distinct()
    Logical: any(), all()

# `pull()`

Sometimes you want to extract a single column from a data frame as a *vector* (or single value). 

 *pulls* a column of a data frame out as a vector.

```{r}
yugoslavia %>%
    filter(year >= 1982, country == "Serbia") #%>% 
    #summarize(mean_life_exp = mean(lifeExp)) #%>% 
    #summarize( bigest_population = max(pop)) %>%
    #pull()

```


For the year 1982, let's get the number of observations, total population, mean life expectancy, and range of life expectancy for former Yugoslavian countries.

```{r}
yugoslavia %>%
    filter(year == 1982) %>%
    summarize(n_obs          = n(),
              total_pop      = sum(pop),
              mean_life_exp  = mean(lifeExp),
              range_life_exp = max(lifeExp) - min(lifeExp)) 

```
# `group_by()`


The special function `group_by()` changes how functions operate on the data, most importantly `summarize()`.

Functions after `group_by()` are computed *within each group* as defined by variables given, rather than over all rows at once. Typically the variables you group by will be integers, factors, or characters, and not continuous real values.


```{r}
yugoslavia %>%
  group_by(year) %>% 
    summarize(num_countries     = n_distinct(country),
              total_pop         = sum(pop),
              total_gdp_per_cap = sum(pop*gdpPercap)/total_pop) 

```

```{r}
gapminder %>%
  group_by(year) %>% 
    summarize(num_countries     = n_distinct(country),
              total_pop         = sum(pop),
              total_gdp_per_cap = sum(pop*gdpPercap)/total_pop) %>% 
  ggplot(aes(x= year, y = total_pop))+
  geom_line() + geom_point()
  
```


```{r}
gapminder %>%
    #filter(continent != "Asia") %>% 
    group_by(continent, year) %>%
       summarize(num_countries     = n_distinct(country),
              total_pop         = sum(pop),
              total_gdp_per_cap = sum(pop*gdpPercap)/total_pop) %>% 
    ggplot(., aes(x=year, y=total_pop, color=continent)) +
      geom_point(size=1.5) +
      geom_smooth(aes(fill=continent), method="lm")
```

Polt life expectancy in Canada through years compared to US and UK:
```{r}
gapminder %>% 
          filter(country %in% c("Canada","United States", "United Kingdom")) %>% 
          ggplot(aes (x= year , y = lifeExp))+
              geom_line( aes(color= country) )
```
Polt life expectancy vs GDP per capita in 1952 and 2007 for Canada, US and UK:

```{r}
gapminder %>% 
        filter(year %in% c(1952 , 2007)) %>%
        filter(country %in% c("Canada","United States", "United Kingdom") ) %>%
        ggplot(aes(gdpPercap, lifeExp, col = country)) +
        geom_point() +
        facet_grid(. ~ year)

```

 
Which country in the Americas had the shortest life expectancy in months in the year 1962? 

```{r}
gapminder %>%
    mutate(lifeExpMonths = 12 * lifeExp) %>%
    filter(year == 1962, continent =='Americas') %>%
    arrange(lifeExpMonths)
```

