---
title: "Review some topics part1"
author: "Mohsen Bahremani"
date: "10/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Gapminder Data

We will be working with data from Hans Rosling's [Gapminder](http://www.gapminder.org) project.
An excerpt of these data can be accessed through an R package called `gapminder`, cleaned and assembled by Jenny Bryan at UBC.

--

If you did not install Gapminder, Please run this in the console: `install.packages("gapminder")`

Load the package and data:
```{r}
library(gapminder)
library(tidyverse)
```

# Initial step

It has been suggested that at initial step, we look over the data, including:

Number of variables,
Number of observations, 
definition of varialbes, 
and so on.
```{r}
glimpse(gapminder)
```

# Interesting point is that:

* **Factor** variables `country` and `continent`

   + Factors are categorical data with an underlying numeric representation

--

* For number of observation, you can use nrow(gapminder) in addition of the above output. 

--

* A nested/hierarchical structure: `year` in `country` in `continent`

   + These are panel data!
   
--
# Description of the Data

```{r}
?gapminder
```

Also, if you use external datasets like some CSV files, this code , i.e. "attributes(NAME_OF_DATA)" or especially for description of variables attributes(NAME_OF_DATA)$dictionary, can be useful.

--
# Summary

General information about:
  + the statistics of variables 
  + the number of rows related to factor variables 


```{r}
summary(gapminder)
```


# plots 

-- Bar chart
   Bar plots are often used to visualize the distribution of a discrete variable, like continent. 
   With ggplot2, this is relatively easy:

    map the x variable to continent
    add a geom_bar() layer, that counts the observations in each category and plots them as bar lengths.



```{r}
ggplot(gapminder, aes(x=continent, color=continent)) + 
  geom_bar()

# How the colore of whole of the rectangle can be changed?
```

one of the method to store the plot in order to call later:

```{r}
mybar <- last_plot()
```

we can add layer to the stored plot

```{r}
mybar + coord_flip()
mybar + coord_polar()
```

-- boxplot
Alternatively, you might want to view the distributions of life expectancy by another visual summary, grouped by continent.
All you need to do is change the aesthetic to show continent on one axis, and life expectancy (lifeExp) on the other.
```{r}
ggplot(data=gapminder, aes(x=continent, y=lifeExp, fill=continent))+
        geom_boxplot(outlier.size=2)
        
```

-- scatter plot
    Now let’s explore the relationship between life expectancy and GDP (Gross Domestic Product) with a scatterplot
```{r}
ggplot(gapminder, aes(x = gdpPercap, y = lifeExp))+
  geom_point()

# how we can do more readable this plot?
```

-- 
GDP variable is so skewed that it is better plotted on a log scale:

```{r}
ggplot(gapminder, aes(x = log10(gdpPercap), y = lifeExp)) +
  geom_point() #+
  #xlab(" Logarithm of GDP") + 
  #ylab("Life expectancy") 
# improvement?
```


Changing theme and font size


# Plotting All Countries

```{r}
ggplot(data = gapminder, 
       aes(x = year, y = lifeExp)) +
  geom_point(color = "red", size = 3) +
  #geom_line(color = "red", size = 3) +
  xlab("Year") + 
  ylab("Life expectancy") +
  ggtitle("Life expectancy over time") +
  theme_bw(base_size=18)
```

# simple grouping

```{r}
ggplot(data = gapminder, 
       aes(x = year, y = lifeExp, 
           group = country)) +
  geom_line(color = "red") + #<<
  xlab("Year") + 
  ylab("Life expectancy") +
  ggtitle("Life expectancy over time") +
  theme_bw(base_size=18)
```


```{r}
ggplot(data = gapminder, 
       aes(x = year, y = lifeExp, 
           group = country, 
           color = continent)) + 
  geom_line() + 
  xlab("Year") + 
  ylab("Life expectancy") +
  ggtitle("Life expectancy over time") +
  theme_bw(base_size=18)
```
 Seperating the continet plot and modification on legend position
```{r}
ggplot(data = gapminder, 
       aes(x = year, y = lifeExp, 
           group = country, 
           color = continent)) +
  geom_line() +
  xlab("Year") + 
  ylab("Life expectancy") +
  ggtitle("Life expectancy over time") +
  theme_bw(base_size=18) +
  facet_wrap(~ continent)
  #theme(legend.position = c(0.8, 0.25))
  #theme(legend.position = "none")
```
# Showing a Stored Graph

```{r}
lifeExp_by_year <- last_plot()
```

--call the stored graph
--add layer to stored graph

```{r}
lifeExp_by_year #+
  theme(legend.position = "bottom")
```


```{r}

lifeExp_by_year +
  theme(legend.position = c(0.8, 0.25))+
  scale_color_manual(
      name = "Which\ncontinent\nare we\nlooking at?", # \n adds a line break #<<
      values = c("Africa" = "seagreen", "Americas" = "turquoise1", 
                 "Asia" = "royalblue", "Europe" = "violetred1", "Oceania" = "yellow"))
```
 

