---
title: "Review some topics part1"
author: "Mohsen Bahremani"
date: "10/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Gapminder Data

We will be working with data from Hans Rosling's [Gapminder](http://www.gapminder.org) project.
An excerpt of these data can be accessed through an R package called `gapminder`, cleaned and assembled by Jenny Bryan at UBC.

--

If you did not install Gapminder, Please run this in the console: `install.packages("gapminder")`

Load the package and data:
```{r}
library(gapminder)
library(tidyverse)
```

# Initial step

It has been suggested that at initial step, we look over the data, including:

Number of variables,
Number of observations, 
definition of varialbes, 
and so on.
```{r}
glimpse(gapminder)
```

# Interesting point is that:

* **Factor** variables `country` and `continent`

   + Factors are categorical data with an underlying numeric representation

--

* For number of observation, you can use nrow(gapminder) in addition of the above output. 

--

* A nested/hierarchical structure: `year` in `country` in `continent`

   + These are panel data!
   
--
# Description of the Data

```{r}
?gapminder
```

Also, if you use external datasets like some CSV files, this code , i.e. "arrtibutre(NAME_OF_DATA)" or especially for description of variables arrtibutre(NAME_OF_DATA)$dictionary, can be useful.

--
# Summary

General information about:
  + the statistics of variables 
  + the number of rows related to factor variables 


```{r}
summary(gapminder)
```
# `filter` Data Frames

How we can choose on country like Brazil 
```{r}
gapminder %>% filter(country == "Brazil")
```


We used `==` for testing "equals": `country == "Brazil"`.

--

* `!=`: not equal to

* `>`, `>=`, `<`, `<=`: less than, less than or equal to, etc.

* `%in%`: used with checking equal to one of several values

--

Or we can combine multiple logical conditions:

* `&`: both conditions need to hold (AND)

* `|`: at least one condition needs to hold (OR)

* `!`: inverts a logical condition (`TRUE` becomes `FALSE`, `FALSE` becomes `TRUE`)

--


# Examples

 observations from Brazil after 1990 and through 2000.
```{r}
gapminder %>%
    filter(country == "Brazil" &
           year > 1990 &
           year <= 2000 )
```

two first of observations from China who were expect to live more than 60 years  after 1950.

```{r}
China <- gapminder %>% 
          filter(country == "China" &
                lifeExp > 60 &
                year > 1950)
head(China, 2)
```

# plots 

Bar chart
```{r}
ggplot(gapminder, aes(x=continent, color=continent)) + 
  geom_bar()

ggplot(gapminder, aes(x=continent, fill=continent)) + 
  geom_bar()
```

```{r}
mybar <- last_plot()
```

```{r}
mybar + coord_flip()
mybar + coord_polar()
```


```{r}
ggplot(data=gapminder, aes(x=continent, y=lifeExp, fill=continent))+
        geom_boxplot(outlier.size=2)
        
```

```{r}
ggplot(gapminder, aes(x = gdpPercap, y = lifeExp))+
  geom_point()
```

```{r}
ggplot(gapminder, aes(x = gdpPercap, y = lifeExp))+
       geom_point(aes(color=continent)) #+
       #geom_smooth() 
```

```{r}
ggplot(gapminder, aes(x = log10(gdpPercap), y = lifeExp)) +
  geom_point()
```

```{r}
 ggplot(gapminder, aes(x = gdpPercap, y = lifeExp)) +
        scale_x_log10() +
        geom_point(aes(color=continent)) +
        geom_smooth()
```


Ploting Chinese Life Expectancy  

```{r}
ggplot(data = China, 
       aes(x = year, y = lifeExp)) +
    geom_point()
```




simple modification 

```{r}
ggplot(data = China, 
       aes(x = year, y = lifeExp)) +
  geom_point(color = "red", size = 3)
```

Changing the axis' and plot's label 

```{r}
ggplot(data = China, 
       aes(x = year, y = lifeExp)) +
  geom_point(color = "red", size = 3) +
  xlab("Year") +
  ylab("Life expectancy")+
  ggtitle("Life expectancy in China")
```
Changing theme and font size

```{r}
ggplot(data = China, 
       aes(x = year, y = lifeExp)) +
  geom_point(color = "red", size = 3) +
  #geom_line(color = "red", size = 3) +
  xlab("Year") +
  ylab("Life expectancy")+
  ggtitle("Life expectancy in China")+
  theme_bw()
  #theme_bw(base_size=18) 
```
# Plotting All Countries

```{r}
ggplot(data = gapminder, 
       aes(x = year, y = lifeExp)) +
  geom_point(color = "red", size = 3) +
  #geom_line(color = "red", size = 3) +
  xlab("Year") + 
  ylab("Life expectancy") +
  ggtitle("Life expectancy over time") +
  theme_bw(base_size=18)
```

# simple grouping 
```{r}
ggplot(data = gapminder, 
       aes(x = year, y = lifeExp, 
           group = country)) +
  geom_line(color = "red") + #<<
  xlab("Year") + 
  ylab("Life expectancy") +
  ggtitle("Life expectancy over time") +
  theme_bw(base_size=18)
```


```{r}
ggplot(data = gapminder, 
       aes(x = year, y = lifeExp, 
           group = country, 
           color = continent)) + #<<
  geom_line() + #<<
  xlab("Year") + 
  ylab("Life expectancy") +
  ggtitle("Life expectancy over time") +
  theme_bw(base_size=18)
```
 Seperating the continet plot and modification on legend position
```{r}
ggplot(data = gapminder, 
       aes(x = year, y = lifeExp, 
           group = country, 
           color = continent)) +
  geom_line() +
  xlab("Year") + 
  ylab("Life expectancy") +
  ggtitle("Life expectancy over time") +
  theme_bw(base_size=18) +
  facet_wrap(~ continent)
  #theme(legend.position = c(0.8, 0.25))
  #theme(legend.position = "none")
```
# Showing a Stored Graph

```{r}
lifeExp_by_year <- 
  ggplot(data = gapminder, 
       aes(x = year, y = lifeExp, 
           group = country, 
           color = continent)) +
  geom_line() +
  xlab("Year") + 
  ylab("Life expectancy") +
  ggtitle("Life expectancy over time") +
  theme_bw() + 
  facet_wrap(~ continent) +
  theme(legend.position = "none")
```

--call the stored graph
--add layer to stored graph

```{r}
lifeExp_by_year #+
  theme(legend.position = "bottom")
```


```{r}

lifeExp_by_year +
  theme(legend.position = c(0.8, 0.25))+
  scale_color_manual(
      name = "Which\ncontinent\nare we\nlooking at?", # \n adds a line break #<<
      values = c("Africa" = "seagreen", "Americas" = "turquoise1", 
                 "Asia" = "royalblue", "Europe" = "violetred1", "Oceania" = "yellow"))
```
 

